---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

```{r echo=TRUE}
rm(list=ls())
setwd("/Users/wenjian/Desktop/R programming/Reproducible Research/Assignment 1/RepData_PeerAssessment1/")
library(data.table)
data<-fread("activity.csv")
```

## What is mean total number of steps taken per day?
```{r echo=TRUE}
result1<-data[,.(meansteps=mean(steps),totalsteps=sum(steps)),by=.(date)]
hist(result1$totalsteps,main="Histogram of total number of steps",xlab="Total steps each day", ylab="Number of days")
library(knitr)
kable(result1,col.names = c("Dates","Mean steps per day","Total steps per day"))
```

## What is the average daily activity pattern?
```{r echo=TRUE}
data1<-as.data.frame(data)
data1[is.na(data1)]<-0    #remove NAs
data1<-as.data.table(data1)
result2<-data1[,.(meansteps=mean(steps)),by=.(interval)]
plot(result2$interval,result2$meansteps,type="l",xlab="Intervals",ylab="Mean number of steps",main="Average daily activity pattern")
result2<-result2[order(-meansteps)]
kable(head(result2,5),caption = "Top 5 intervals with highest activities", col.names = c("Top 5 intervals","Mean number of steps"))
```

The interval with the highest mean number of steps is interval `r result2$interval[1]`.

## Imputing missing values
```{r echo=TRUE}
missingrow<-data[is.na(steps),.N]
```

1) There are `r missingrow` number of missing rows.
2) Strategy for imputing missing numbers would be to replace NAs with the **mean for each interval**

```{r echo=TRUE}
data2<-fread("activity.csv")
count=0
result2$meansteps<-as.integer(result2$meansteps)
for(i in 1:nrow(data2)){
        if(is.na(data2[i]$steps)){
                data2[i]$steps<-result2$meansteps[match(data2[i]$interval,result2$interval)]
                count=count+1
        }
}

cat("Total NA replaced is",count)

result3<-data2[,.(meansteps=mean(steps),totalsteps=sum(steps)),by=.(date)]
hist(result3$totalsteps,main="Histogram of total number of steps",xlab="Total steps each day", ylab="Number of days")
meanmedtotal1<-result1[!is.na(totalsteps),.(mean=mean(totalsteps),med=median(totalsteps))]
meanmedtotal2<-result3[,.(mean=mean(totalsteps),med=median(totalsteps))]

kable(meanmedtotal1,caption ="With NAs in data",col.names = c("Mean of total steps","Median of total steps"))
kable(meanmedtotal2,caption="With NAs replaced",col.names = c("Mean of total steps","Median of total steps"))
```

## Are there differences in activity patterns between weekdays and weekends?

```{r echo=TRUE}
library(lubridate)
data2$date<-ymd(data2$date)
data2$weekday<-"factors" #Create a new column for ID-ing weekdays or weekends
    
for(i in 1:nrow(data2)){
                if(weekdays(data2[i]$date)=="Saturday" | weekdays(data2[i]$date)=="Sunday"){
                data2[i]$weekday<-"Weekend"
        } 
                else(data2[i]$weekday<-"Weekday") 
        }

resultwd<-data2[,.(meansteps=mean(steps)),by=.(interval,weekday)]

library(ggplot2)
ggplot(resultwd,aes(x=interval,y=meansteps))+
        geom_line()+
        facet_wrap(~weekday)+
        xlab("Interval")+
        ylab("Number of steps")+
        ggtitle("Difference between weekday and weekends")


```